Amendments Summary

### 2025-01-02 Thesis/amend_v2.qmd
amendment summary

## Responses to Comments from Examiner Prof. Louise Ryan:

@@

Thank you for your valuable feedback on the presentation and accessibility of the abstract. I appreciate your suggestions to make the abstract more engaging and suitable for a non-expert audience.

I have revised the abstract to ensure that it provides a clear and concise overview of the thesis in a manner accessible to a broader audience, including non-experts. Technical terms such as "lineup protocol" have been either omitted or explained briefly with contextual statements to ensure clarity for readers unfamiliar with these concepts. For instance, I have included a sentence explaining that the lineup protocol is a statistical visualization approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.

@@

Thank you for this suggestion. I have added a statement in Chapter 1 to highlight the wide-ranging applications of regression diagnostics across various models. I also clarified that the primary focus of my thesis is on diagnostics for classical normal linear regression model to establish this context early on.

@@

I agree that including a visual representation of the lineup protocol in the introduction would help clarify its concept and engage the reader. I have added a diagram in Chapter 1 to illustrate how the lineup protocol works, providing an example similar to those discussed in Chapter 2. This addition should help familiarize readers with the protocol from the outset.

@@

The recruitment details are provided in the appendix. The subjects received training by reading a webpage with examples. Unfortunately, I am unable to modify this chapter further, as it has already been submitted as a paper and accepted, requiring it to remain unchanged.

@@

I completely agree that Tukey’s contributions were foundational and greatly impactful. While I appreciate the suggestion to add more references to his work earlier in the thesis, I did not find a suitable place where they would naturally fit within the context. However, I’m glad the reference in Chapter 3 resonated with you.

@@

I appreciate your careful observation. I’ve addressed the issue by improving the language for better clarity in Chapter 3, such as changing "dynamic input size" to "variable input size" and providing a clear definition on page 30. I appreciate your attention to detail and constructive feedback!

@@

Yes, the actual plot and all the null plots for the lineup protocol are generated prior to being input into the computer vision model. The chapter primarily focuses on residuals vs. fitted values plots. Other types of plots, such as residuals vs. specific predictor variables, are discussed in Section 3.12, which covers limitations and future work.

@@

Thank you for pointing this out. To avoid confusion, I have clarified the terminology by changing "model specification" in Section 3.2 to "Computer Vision Model Specifications." This should make it clearer that the term refers specifically to the details of the computer vision model and not the underlying statistical model being critiqued. I appreciate your feedback!

@@

I appreciate you bringing this to my attention. I have added a paragraph in Section 3.7 to provide more details on how the predictors $x_1$ and $x_2$ in the training datasets are generated. I appreciate your careful review and feedback!

@@

I’m grateful for your observation. The placement of Figure 3.1 was indeed a formatting issue in the original document. This has since been fixed, and the figure now appears much closer to its citation in the text. I appreciate your attention to this detail!

## Responses to Comments from Examiner Prof. Daniela Witten:

@@

Thank you for the insightful question. This is something I considered carefully and discussed extensively with my supervisors during the development of the methodology. While training a deep learning model to predict categories like "contains outliers" vs. "does not contain outliers" or "is non-linear" vs. "not non-linear" seems appealing, it would bring us back to the traditional testing framework. These tests typically focus on detecting specific types of model violations in isolation.

One of the key advantages of using the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided those violations produce noticeable differences in the data plots. In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.

Additionally, if the ultimate goal is to distinguish between specific cases (e.g., outliers vs. no outliers, or linear vs. non-linear relationships), conventional statistical tests based on regression model assumptions are often more suitable. Computer vision models may not be the best tools for these tasks. Even if we trained a deep learning model for this purpose, it would still face challenges in terms of confidence assessment. Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.

The lineup evaluation described in Section 3.5 remains valuable even in this context. By treating the model output as a statistic and comparing it to outputs generated under null assumptions, it adheres to the principles of visual inference and conventional statistical testing. This approach provides a robust framework for interpreting results and avoids the pitfalls of relying solely on predicted probabilities.

@@

In contrast, computer vision models can evaluate lineups with as many null plots as desired, which is one of their advantages.

@@

Thank you for this valuable suggestion. I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal.

@@

>>>>

## Changes made in response to comments from Examiner Prof. Louise Ryan

@@

<!--Thank you for your valuable feedback on the presentation and accessibility of the abstract. I appreciate your suggestions to make the abstract more engaging and suitable for a non-expert audience.-->

I have revised the abstract to ensure that it provides a clear and concise overview of the thesis in a manner accessible to a broader audience, including non-experts. Technical terms have been either omitted or explained briefly with contextual statements to ensure clarity for readers unfamiliar with these concepts. For example, I have included a sentence explaining that the lineup protocol is a statistical visualisation approach that embeds a data plot among null plots to compare patterns arising by chance with those observed in the data.

@@

I have added a statement in Chapter 1 to highlight the wide-ranging applications of regression diagnostics across various models. I also clarified that the primary focus of my thesis is on diagnostics for classical normal linear regression model to establish this context early on.

@@

I have added a diagram in Chapter 1 to illustrate how the lineup protocol works, providing an example similar to those discussed in Chapter 2. This addition helps clarify its concept and engage the reader with the protocol from the outset.

@@

The recruitment details are provided in the appendix. The subjects received training by reading a webpage with examples. I decided not to modify this chapter further, as it has been submitted published.

@@

I completely agree that Tukey’s contributions were foundational and greatly impactful. While I appreciate the suggestion to add more references to his work earlier in the thesis, this one is sufficient to point to that body of work, and accommodate references to the range of foundational research on regression analysis.

@@

I’ve addressed the issue by generally improving the language for better clarity in Chapter 3, including changing "dynamic input size" to "variable input size" and providing a clear definition on page 30.

@@

Yes, the actual plot and all the null plots for the lineup protocol are generated prior to being input into the computer vision model. The chapter primarily focuses on residuals vs. fitted values plots. Other types of plots, such as residuals vs. specific predictor variables, are discussed in Section 3.12, which covers limitations and future work. No changes were made.

@@

To avoid confusion, I have clarified the terminology by changing "model specification" in Section 3.2 to "computer vision model specification" This should make it clearer that the term refers specifically to the details of the computer vision model and not the underlying statistical model being critiqued.

@@

I have added a paragraph in Section 3.7 to provide more details on how the predictors $x_1$ and $x_2$ in the training datasets are generated.

@@

The placement of Figure 3.1 was indeed a formatting issue in the original document. This has since been fixed, and the figure now appears much closer to its citation in the text.

## Changes made in response to comments from Examiner Prof. Daniela Witten

@@

This is something I considered carefully and discussed extensively with my supervisors during the development of the methodology. While training a deep learning model to predict categories like "contains outliers" vs. "does not contain outliers" or "is non-linear" vs. "not non-linear" seems appealing, it would bring us back to the traditional testing framework. These tests typically focus on detecting specific types of model violations in isolation.

One of the key advantages of using the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided those violations produce noticeable differences in the data plots. <!--In contrast, training a model discussed above would require us to explicitly define and map out all possible violations of interest—or at least the common ones we want to address.

Additionally, if the ultimate goal is to distinguish between specific cases (e.g., outliers vs. no outliers, or linear vs. non-linear relationships), conventional statistical tests based on regression model assumptions are often more suitable. Computer vision models may not be the best tools for these tasks. Even if we trained a deep learning model for this purpose, it would still face challenges in terms of confidence assessment. Such a model would typically produce a single probability prediction for each predefined violation, limiting its scope and flexibility.-->

The lineup evaluation described in Section 3.5 remains valuable even in this context. By treating the model output as a statistic and comparing it to outputs generated under null assumptions, it adheres to the principles of visual inference and conventional statistical testing. This approach provides a robust framework for interpreting results and avoids the pitfalls of relying solely on predicted probabilities. **XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela? If so, this needs to be stated here.**

@@

In contrast, computer vision models can evaluate lineups with as many null plots as desired, which is one of their advantages. **XXX Have you made any modifications in the thesis to make these points so that the reader is not likely to ask the same question as Daniela? If so, this needs to be stated here.**

@@

I will include a discussion on this point in the manuscript currently being prepared for the peer-reviewed journal. **XXX I think you need to add a sentence like this to the thesis too.**

@@

**XXX Not sure if this is a good answer here. The main purpose of avoiding the human evaluation is to reduce the work required of humans. So while it is a potentially good idea to have this hybrid approach the problem is that it won't be adopted because it is too time-consuming. traditional tests are often done because its a click on a software interface or a simple line of code. While making a lineup is typically a simple line of code, too, it requires the additional work of staring at a page of plots to assess if one can determine a difference. So I don't think that you need to respond in the way you have done. But you do need to state how you modified your thesis or why not modify it.**
